\input{header}
\input{macros}
\sheet{10}{January 10th, 2025}
\setcounter{section}{1}
\begin{document}

\section{Nonregular languages}

\subsection{(a)}
$$A = \left\{ 0^{3^{n}} \mid n \in \NN \right\}$$
We use the Pumping Lemma.\\
Consider $a \in A$, $a = uvw$. $u=\epsilon, v=0^{3^{n}}, w=\epsilon$ for a $n \in \NN$.\\
We decompose $v$ into $v=xyz$, where $x=y=z=0^{3^{n-1}}$. Specifically, $y\neq\epsilon$.\\
We now consider $uxy^{i}zw = \epsilon 0^{3^{n-1}} ( 0^{3^{n-1}})^{i} 0^{3^{n-1}} \epsilon$.\\
Setting $i=2$, we get: $uxy^{2}zw=0^{3^{n-1}}0^{2\cdot3^{n-1}}0^{3^{n-1}}$.\\
Now, we want to show that this is not part of $A$. For this, we show that $3^{n-1}+2\cdot 3^{n-1} + 3^{n-1} = 4\cdot 3^{n-1}$ is not a power of $3$.\\
$4\cdot 3^{n-1} > 3^{n-1}$ obviously.\\
$4\cdot 3^{n-1} > 3\cdot 3^{n-1} = 3^{n}$\\
$4\cdot 3^{n-1} < 9\cdot 3^{n-1} = 3^{2}\cdot 3^{n-1} = 3^{n+1}$\\
We have shown that $4\cdot 3^{n-1}$ is larger than one power of $3$, but smaller than the next. Therefore, it cannot be equal to any power of $3$.\\
This shows that $A$ does not have the pumping property. Therefore, $A$ is not regular.

\subsection{(b)}
$$B = \left\{ (a^{k}bc^{j})^{n} \mid k,j \in \NN\text{ and }n\text{ is not prime} \right\}$$
We use the Pumping Lemma.\\
Consider $b=(a^{k}bc^{j})^{n} = uvw$.\\
We decompose this into $u=a^{k}bc^{j}$, $v=a^{k}bc^{j}$, $w=(a^{k}bc^{j})^{n-2}$.\\
Now we consider an arbitrary decomposition $v=xyz$ with $y\neq \epsilon$.\\
We do a case distinction on the value of $v$:
\begin{itemize}
    \item $v= b$\\
    In this case, with $i=2$, we get $uxy^{i}zw = a^{k}bc^{j}a^{k}b^{2}c^{j}(a^{k}bc^{j})^{n-2}$. This contains 2 successive $b$ characters, which cannot occur in our language, therefore $uxy^{i}zw \notin B$.
    \item $v= a^{m}b$ or $v= "bc^{n}"$ with $m \leq k$ and $n \leq j$\\
    Here, with $i=2$, we get a succession of $a^{m}b$ with no $c$ characters or a succession of $bc^{n}$ with no $a$ characters, which doesn't fulfill the $a^{k}bc^{j}$ pattern. Therefore, $uxy^{i}zw \notin B$.
    \item $v= a^{m}bc^{n}$ with $m \leq k$ and $n \leq j$\\
    Here, with $i=2$, we get $v=a^{k-m}(a^{m}bc^{n})(a^{m}bc^{n})b^{j-n}$. This does not fit the pattern $a^{k}bc^{j}$
    \item $v= a^{k}bc^{j}$\\
    Here, we get $uxy^{i}zw = (a^{k}bc^{j})^{n+i-1}$.\\
    We know that there are infinitely many prime numbers, so we know: $\forall n \in \NN: \exists k \in \NN: n+k \text{ is prime}$.\\
    By selecting $i$ appropriately, we can construct this such that $n+i-1$ is prime, such that $uxy^{i}zw = (a^{k}bc^{j})^{n+i-1} \notin B$
\end{itemize}
Therefore, $B$ does not fulfill the pumping property and is therefore not regular.


\subsection{(c)}
$$C = \left\{ 0^{i}1^{j}2^{k} \mid i\cdot j \neq k \right\}$$
We use the Myhill-Nerode Theorem.\\
Consider the equivalence classes $x_{n} = 0^{n}1^{n}$ and $x_{m} = 0^{n}1^{m}$ where $n\neq m$.\\
Append $2^{n^{2}}$ to both:
\begin{itemize}
    \item $0^{n}1^{n}2^{n^{2}} \notin C$ since $n \cdot n = n^{2}$
    \item $0^{n}1^{m}2^{n^{2}} \in C$ since $n \cdot m \neq n^{2}$ because $n \neq m$
\end{itemize}
We have shown that the equivalence classes $x_{n}$ and $x_{m}$ are pairwise different for and $n,m \in \NN$. This shows that there are infinitely many Myhill-Nerode equivalence classes- Therefore, according to the Myhill-Nerode Theorem, $C$ is not regular.


\subsection{(d)}
$$D = \left\{ x \in \left\{(,)\right\}^{\ast} \mid x \text{ is a balanced string of parentheses} \right\}$$
We use the Myhill-Nerode Theorem.\\
Consider the equivalence classes $x_{n} = (^{n}$ ($n$ opening parentheses), and $x_{m} = (^{m}$ ($m$ opening parentheses), with $n>m$.\\
We now append $n$ closing parentheses to both:\\
\begin{itemize}
    \item $(^{n})^{n} \in L$ since it is balanced with both $n$ opening and closing parentheses
    \item $(^{m})^{n} \notin L$ since there are more closing than opening parentheses (since $n>m$)
\end{itemize}
We have shown that $x_{n} \nsim x_{m}$, which implies that there are infinitely many Myhill-Nerode equivalence classes. Therefore, the language $D$ is not regular.

\end{document}
